Multi-target detection with application to cryo-electron microscopy

### 2019-03-04 inverse_problems/inverse_problems.tex
section 2

Even if particle picking is feasible, it may suffer from a phenomenon called \emph{model bias}.
The main mechanism leads most particle pickers is based on correlating the micrograph with several templates; areas with high correlation are identified as areas that  are likely to contain a projection (these are the picked particles). 
Clearly, the picked particles depend heavily on the chosen templates.
The potential pitfall of model biased
has been recognized early by the cryo-EM community~\cite{shatsky2009method,vanheel1992correlation,henderson2013avoiding,vanheel2013finding}. In the classical ``Einstein from noise'' experiment, multiple realizations of pure noise are aligned to a picture of Einstein using template matching and then averaged. It was shown that the averaged noise rapidly becomes remarkably similar to the Einstein template~\cite{shatsky2009method}. In the context of cryo--EM, this experiment exemplifies how prior assumptions about the particles biases the reconstructed structure. %This model bias is common to all particle picking methods based on template matching. %In our approach, no templates or human intervention are required, thus significantly reducing concerns about model bias. %

A recent work of the authors suggests to bypass the particle picking stage and  reconstruct the 3-D structure directly from the micrograph~\cite{bendory2018toward}.
Based on autocorrelation analysis it was shown that---at least in principle---the limits particle picking imposes on molecule size do not necessarily  translate into limits on particle reconstruction.
%The principle mathematical tool is  \emph{autocorrelation analysis}, described in detail in Section~\ref{sec:AC_analysis}.
This goal of  the current paper is to provide a theoretical justification and numerical support  for the method proposed in~\cite{bendory2018toward}.  
%In this context, the  models described above serve  as  an abstraction of the cryo-EM problem: the random signal $X$ discussed in Section~\ref{sec:extensions} can be thought of as 2-D random tomographic projections of the 3-D structure 
%taken according to some unknown distribution of the particles within the ice.

% A similar abstraction has been proven itself to be useful in a similar context as described in the next section.

%We mention that~\cite{bendory2018toward} was not the first paper to employ autocorrelation analysis to cryo-EM. 
Autocorrelation analysis for cryo-EM traces back to the seminal paper of Zvi Kam~\cite{kam1980reconstruction}. This paper proposed autocorrelation analysis for {3-D} reconstruction, under the assumption of picked, perfectly centered, particles. Kam's method has been extended and employed in X-ray free electron lasers and cryo-EM~\cite{liu2013three,kurta2017correlations,levin20173d,von2018structure}.  
In order to investigate the computational and statistical properties of Kam's method, a series of papers have studied a simplified model, 
called   \emph{multi-reference alignment}~\cite{bandeira2014multireference,bendory2017bispectrum,bandeira2017optimal,perry2017sample,bandeira2017estimation,abbe2017multireference}.
We follows the same line of research by considering the multi-target detection as an abstraction to the application of reconstructing 3-D structures directly from the micrograph as proposed in~\cite{bendory2018toward}.
\TODO{Not sure about the location of the last paragraph.}

>>>>

Even if particle picking is feasible, it may suffer from a phenomenon called \emph{model bias}.
The main mechanism guides  most particle pickers is based on correlating the micrograph with several templates; highly correlated areas are associated with projections (these are the picked particles). 
Clearly, the picked particles depend heavily on the chosen templates: different templates may lead to different picked particles, and thus to different 3-D structures.
This potential pitfall 
has been recognized early by the cryo-EM community~\cite{vanheel1992correlation,henderson2013avoiding,vanheel2013finding} and was  exemplified using the
``Einstein from noise'' experiment~\cite{shatsky2009method}. %In this experiment, multiple realizations of pure noise are aligned to a picture of Einstein using template matching and then averaged. It was shown that the averaged noise rapidly becomes remarkably similar to the Einstein template.
 %In the context of cryo--EM, this experiment exemplifies how prior assumptions about the particles biases the reconstructed structure. %This model bias is common to all particle picking methods based on template matching. %In our approach, no templates or human intervention are required, thus significantly reducing concerns about model bias. %

A recent work of the authors suggests a methodology to bypass particle picking and  reconstruct the 3-D structure directly from the micrograph~\cite{bendory2018toward}.
Based on autocorrelation analysis, it was shown that---at least in principle---the limits high noise regimes 
impose on particle picking  do not necessarily translate into limits on 3-D reconstruction.
The main goal of  this paper is to provide theoretical justification and numerical support  for this method.

In the next section, we introduce autocorrelation analysis in detail, focusing on the multi-target detection model. 
We mention that similar ideas in  cryo-EM trace back to the seminal paper of Zvi Kam~\cite{kam1980reconstruction}. Kam proposed autocorrelation analysis for {3-D} reconstruction, under the assumption of picked, perfectly centered, particles. Kam's method has been extended and employed in X-ray free electron lasers and cryo-EM~\cite{liu2013three,kurta2017correlations,levin20173d,von2018structure}.  
In order to investigate the computational and statistical properties of Kam's method, a series of papers have studied a simplified model, 
called   \emph{multi-reference alignment}~\cite{bandeira2014multireference,bendory2017bispectrum,bandeira2017optimal,perry2017sample,bandeira2017estimation,abbe2017multireference}.
We follows the same line of research by considering the multi-target detection as an abstraction to the application of reconstructing 3-D structures directly from the micrograph~\cite{bendory2018toward}.
